Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-09 manuscript/manuscript.tex
edits

% TODO
% - we have inconsistent notation between intro of results and figure 1 with
%   cov() 
% - address the weird barghi rep correlation point for last time

@@

@@

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the impact of this selection
 genome-wide is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that at least \vb{17\% to 37\%} of allele frequency
 change in these experiments is driven by selection in these experiments.
 Against this background of positive genome-wide temporal covariances we also
 identify signals of negative temporal covariance corresponding to reversals in
 the direction of selection for a reasonable proportion of loci over the time
 course of a selection experiment. Overall, we find that in the three studies
 we analyzed, linked selection has a large impact on short-term allele
 frequency dynamics that is readily distinguishable from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart the
impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of a range of sexual species
\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
time.

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection on genome-wide variation over
short timescales in sexual populations remarkably difficult. This is because
the effect of selection on a polygenic trait (such as fitness) is distributed
across loci in proportion to their effect sizes. This can lead to subtle allele
frequency shifts on standing variation that are difficult to distinguish from
background levels of genetic drift and sampling variance.  However,
increasingly genomic experimental evolution studies with multiple timepoints,
and in some cases multiple replicate populations, are being used to detect
large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from temporal genomic studies is the
temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
These covariances are directly estimable from temporal genomic data and are
created when the loci that underly heritable fitness variation perturb the
frequencies of linked neutral alleles; in contrast, when genetic drift acts
alone in a closed population, these covariances are zero in expectation.
Mathematically, temporal covariances are useful because it is natural to
decompose the total variance in allele frequency change across a set of time
intervals into the variances and covariances in allele frequency change among
time intervals. Furthermore, biologically, these covariances reflect the extent
to which allele frequency changes in one generation predict changes in another
due to a shared genetic pressures.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound on the proportion of total variation in
allele frequency change caused by selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity. The covariances are expected to be zero when only drift is
acting, as only heritable variation for fitness can create covariance between
allele frequency changes in a closed population \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
covariances per replicate are noisier but this general pattern holds; see
Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (e.g. if
adjacent generations had been sequenced) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

\vb{One concern is that these covariances reflect the localized impact of a few
  large-effect loci rather than selection on a polygenic trait. Since our
  covariances are essentially averages over loci, the covariance estimate can
  be strongly affected by a single outlier region. To test whether large
  outlier regions drive the genome-wide signal we see in the
  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
  along the genome (we refer to these as windowed covariances throughout) and
  take the median windowed covariance, and trimmed-mean windowed covariance, as
  a measure of the genome-wide covariance robust to large-effect loci. These
  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean}) confirm
the patterns we see using the mean covariance, confirming that genomic temporal
covariances are non-zero due to the impact of selection on many genomic
windows. }

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure not easily
interpretable. \vb{We can calculate a more intuitive measure from the temporal
  covariances to quantify the impact of selection on allele frequency change:
the ratio of total covariance in allele frequency change to the total variance
in allele frequency change}. We denote the change in allele frequency as
$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
$t$. Since the total variation in allele frequency change can be partitioned
into variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
correct these for sequencing depth), and the covariances are zero when drift
acts alone, this is a lower bound on how much of the variance in allele
frequency change is caused by linked selection \parencite{Buffalo2019-io}. We
call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. Since \textcite{Barghi2019-qy}
experiment is sequenced every ten generations, in the numerator for the
covariance we use the allele frequency changes between adjacent timepoints,
which are ten generations apart. Consequently, this leads our measure $G(t)$ to
be strongly conservative, since the temporal covariances within each
ten-generation block are not not directly observable, and are not included in
the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection.

@@

The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. Intuitively, where temporal covariances
reflect that neutral alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This \vb{suggests that while
a reasonable fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result} consistent with the primary
finding of the original \textcite{Barghi2019-qy} study: that alternative loci
contribute to longer term adaptation \vb{across} the different replicates.

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). Similarly, we can
calculate the proportion of the total variance in allele frequency change
from convergent selection pressure analogous to $G$ where the numerator is is
the convergent covariance and the denominator is the total variance (see
Supplementary Material \ref{supp:replicate-g}). We find that 37\% of the
total variance is due to shared allele frequency changes caused by selection
(95\% CI [29\%, 41\%]; these are similar to the convergence correlation, see
Supplementary Material \ref{supp:replicate-g}).

Next, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
line (Ctrl).  Consequently, this selection experiment offers a useful control
to test our between-replicate covariances: we expect to see positive
between-replicate covariance in the comparison between the two Longshanks
selection lines, but not between the two pairwise comparisons between the
control line and the two Longshanks lines. We find that this is the case (gray
confidence intervals in Figure \ref{fig:figure-2} C), with the two Longshanks
comparisons to the control line not being significantly different from zero,
while the comparison between the two Longshanks lines is statistically
significantly different from zero (CIs [0.013, 0.037]).

One finding in the Longshanks study \vb{was that two} major-effect loci showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that these across-replicate covariances are indeed driven by a large
number of loci. This is consistent with a signal of selection on a polygenic
trait driving genome-wide change, although we note that large-effect loci can
contribute to the indirect change at unlinked loci
\parencite{Robertson1961-ho,Santiago1995-hx}.

>>>>

% - change supp fig ordering
% - sig impact
% -ref table 1
% fix equation G(t) in Fig 1

@@

% Significant Statement:

% Evolutionary biologists and ecologists have repeatedly found evidence of traits
% like flowering time changing rapidly adapting

@@

 Rapid phenotypic adaptation is often observed in natural populations and
 selection experiments.  However, detecting the genome-wide impact of this
 selection is difficult, since adaptation often proceeds from standing
 variation and selection on highly polygenic traits, both of which may leave
 faint genomic signals indistinguishable from a noisy background of genetic
 drift.  One promising signal comes from the genome-wide covariance between
 allele frequency changes observable from temporal genomic data, e.g.
 evolve-and-resequence studies. These temporal covariances reflect how the
 change in neutral allele frequency at one timepoint is predictive of the
 changes at later timepoints when there is heritable fitness variation in the
 population, as neutral alleles can remain associated with selected alleles
 over time. Since genetic drift does not lead to temporal covariance, we can
 use these covariances to estimate what fraction of the variation in allele
 frequency change through time is driven by linked selection. Here, we
 reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
 one artificial selection experiment in mice, to quantify the effects of linked
 selection over short timescales using covariance among time-points and across
 replicates. We estimate that at least 17\% to 37\% of allele frequency
 change \vb{in these populations} is driven by selection in these experiments.
 Against this background of positive genome-wide temporal covariances we also
 identify signals of negative temporal covariance corresponding to reversals in
 the direction of selection for a reasonable proportion of loci over the time
 course of a selection experiment. Overall, we find that in the three studies
 we analyzed, linked selection has a large impact on short-term allele
 frequency dynamics that is readily distinguishable from genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles of
genetic drift and selection in shaping genome-wide allele frequency changes.
Selection can both directly and indirectly affect allele frequencies, with the
indirect effect coming from the action of selection on correlated loci
elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
review). Previous work on this question has mostly focused on teasing apart the
impacts of drift and selection on genome-wide diversity using population
samples from a single contemporary timepoint, often by modeling the correlation
between regional recombination rate, gene density, and diversity created in the
presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
approach has shown linked selection has a major role in shaping patterns of
genome-wide diversity across the genomes of a range of sexual species
\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
However, we lack an understanding of how genome-wide linked selection acts over
time.

There are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
selection-driven genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact of selection on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci in proportion to their effect sizes. This can lead to
subtle allele frequency shifts on standing variation that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
temporal-genomic studies have begun in wild populations, some with the goal of
finding variants that exhibit frequency changes consistent with fluctuating
selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from genomic studies \vb{with multiple
timepoints} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb the frequencies of linked neutral alleles; in
contrast, when genetic drift acts alone in a closed population, these
covariances are zero in expectation.  Mathematically, temporal covariances are
useful because it is natural to decompose the total variance in allele
frequency change across a set of time intervals into the variances and
covariances in allele frequency change among time intervals. Furthermore,
biologically, these covariances reflect the extent to which \vb{neutral} allele
frequency changes in one generation predict changes in another due to a shared
\vb{selection} pressure \vb{and linkage to selected loci}.

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound on the proportion of total variation in
allele frequency change caused by selection, and the correlation between allele
frequency changes between replicate populations caused by \vb{the} response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

\begin{table}
  \centering
  \footnotesize
  \centering
  \begin{tabular}{l c c c c c c}
    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size} & \textbf{Gens.} & \textbf{Timepoints} \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
  \end{tabular}
  \caption{A summary of the main selection studies we analyzed.}
\end{table}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Materials Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). \vb{These}
covariances are expected to be zero when only drift is acting, as only
heritable variation for fitness can create covariance between allele frequency
changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
the ten replicate temporal covariances matrices, we find temporal covariances
that are statistically significant (95\% block bootstraps CIs do not contain
zero), consistent with linked selection perturbing genome-wide allele frequency
changes over very short time periods. The covariances between all adjacent time
intervals are positive and then decay towards zero as we look at more distant
time intervals, as expected when directional selection affects linked variants'
frequency trajectories until ultimately linkage disequilibrium and the additive
genetic variance for fitness associated with neutral alleles decays
\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
but this general pattern holds; see Supplementary Figure
\ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales (e.g. if adjacent generations had been
sequenced) may well be higher yet (see Supplementary Material Section
\ref{supp:barghi-covs} for more details).

One concern is that these covariances reflect the localized impact of a few
large-effect loci rather than selection on a polygenic trait. Since our
covariances are essentially averages over loci, the covariance estimate can be
strongly affected by a single outlier region. To test whether large outlier
regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
data, we calculate the covariances in 100kb windows along the genome (we refer
to these as windowed covariances throughout) and take the median windowed
covariance, and trimmed-mean windowed covariance, as a measure of the
genome-wide covariance robust to large-effect loci. These robust estimates
(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
covariance, confirming that genomic temporal covariances are non-zero due to
the impact of selection \vb{acting across} many genomic windows.

@@

While the presence of positive temporal covariances is consistent with
selection affecting allele frequencies over time, this measure is not easily
interpretable. We can calculate a more intuitive measure from the temporal
  covariances to quantify the impact of selection on allele frequency change:
the ratio of total covariance in allele frequency change to the total variance
in allele frequency change. We denote the change in allele frequency as
$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
$t$. Since the total variation in allele frequency change can be partitioned
into variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
\var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
correct these for sequencing depth), and the covariances are zero when drift
acts alone, this is a lower bound on how much of the variance in allele
frequency change is caused by linked selection \parencite{Buffalo2019-io}. We
call this measure $G(t)$, defined as

\begin{align}
  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
\end{align}
%
which estimates the effect of selection on allele frequency change between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. Since \textcite{Barghi2019-qy}
experiment is sequenced every ten generations, in the numerator for the
covariance we use the allele frequency changes between adjacent timepoints,
which are ten generations apart. Consequently, this leads our measure $G(t)$ to
be strongly conservative, since the temporal covariances within each
ten-generation block are not directly observable, and thus are not included in
the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
than $20\%$ of total, genome-wide allele frequency change over 60 generations
is the result of selection (Figure \ref{fig:figure-1}B).

@@

The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: neutral alleles linked to a particular fitness background are
expected to have allele frequency changes in the same direction if the
selection pressures are similar. Intuitively, where temporal covariances
reflect that neutral alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2}A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2}A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while
a reasonable fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result consistent with the primary
finding of the original \textcite{Barghi2019-qy} study: that alternative loci
contribute to longer term adaptation \vb{across} the different replicates.

@@

A benefit of between-replicate covariances is that unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistically significant
(Figure \ref{fig:figure-2} B, 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859],
[0.00608, 0.00834]). We estimate the convergent correlation coefficient across
these replicates as 0.36 ($95\%$ block-bootstrap confidence interval $[0.31,
0.40]$). Similarly, we can calculate the proportion of the total variance in
allele frequency change from convergent selection pressure analogous to $G$
where the numerator is is the convergent covariance and the denominator is the
total variance (see Supplementary Material \ref{supp:replicate-g}). We find
that 37\% of the total variance is due to shared allele frequency changes
caused by selection (95\% CI [29\%, 41\%]; these are similar to the convergence
correlation, see Supplementary Material \ref{supp:replicate-g}).

Next, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
line (Ctrl).  Consequently, this selection experiment offers a useful control
to test our between-replicate covariances: we expect to see positive
between-replicate covariance in the comparison between the two Longshanks
selection lines, but not between the two pairwise comparisons between the
control line and \vb{each of} the two Longshanks lines. We find that this is
the case (gray confidence intervals in Figure \ref{fig:figure-2} C), with the
two Longshanks comparisons to the control line not being significantly
different from zero, while the comparison between the two Longshanks lines is
statistically significantly different from zero (CIs [0.0129, 0.0400]).

One finding in the Longshanks study \vb{was that two} major-effect loci showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
containing six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2} C),
indicating that these across-replicate covariances are indeed driven by a large
number of loci. This is consistent with a signal of selection on a polygenic
trait driving genome-wide change, although we note that large-effect loci can
contribute to the indirect change at unlinked loci
\parencite{Robertson1961-ho,Santiago1995-hx}.

The presence of an unselected control line provides an alternative way to
partition the effects of linked selection and genetic drift: we can compare the
total variance in allele frequency change of the control line (which excludes
the effect of artificial selection on allele frequencies) to the total variance
in frequency change of the Longshanks selection lines. We can partition the
allele frequency change between the two timepoints (20 generations apart) for a
Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
  p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
  p_{t,\mathrm{LS}}$} where these terms are the decomposition in the allele
  frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
    p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
    p_{t,\mathrm{LS1}}$), and selection response shared between the two
    Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
  similarly for the Longshanks two line, LS2)}. By construction we will assume
  that each of these terms are uncorrelated within replicates, and that only
  the shared term covaries between the replicates.  Assuming that we can
  approximate the contribution of genetic drift in the LS lines as the variance
  in allele frequency change observed in the control, i.e.  \vb{$\var(\Delta
    p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
  p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
  estimate the increase in variance in allele frequency change due to selection
  as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
  \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
  selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
  p_{t,\mathrm{LS2}})$ (see Supplementary Material Section
  \ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
  $[21\%, 48\%]$) of the variance in allele frequency change is driven by the
  effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to
  be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect
  of shared selection between the two Longshanks selection lines.
